\textcolor{violet}{Reverse Correlation Uncovers More Complete Tinnitus Spectra \vspace{0.25cm}}

### 2023-01-05 paper/short/main_wc995.tex
Merge changes from ACL

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
% TC:ignore
\begin{abstract}
  \textit{Goal:} This study validates a reverse correlation method for characterizing the sounds experienced by tinnitus patients with potential for characterizing a wider range of sounds than currently possible.
  \textit{Methods:} Ten normal-hearing subjects assessed the subjective similarity of random auditory stimuli and target tinnitus-like sounds (“buzzing” and “roaring”).
  Reconstructions of the targets were obtained by regressing subject responses on the stimuli, and were compared for accuracy to the frequency spectra of the targets using Pearson's $r$.
  \textit{Results:} Reconstruction accuracy was significantly higher than chance across subjects: buzzing ($M=0.53$, $SD=0.27$): $t(9)=5.766$, $p<0.001$; roaring ($M=0.57$, $SD=0.30$): $t(9)=5.76$, $p<0.001$.
  \textit{Conclusion:} Reverse correlation can accurately reconstruct non-tonal tinnitus-like sounds in normal-hearing subjects,
  indicating its potential for characterizing the sounds experienced by patients with non-tonal tinnitus.
% The purpose of this document is to illustrate how one should prepare manuscripts for submission to the IEEE Open Journal of Engineering in Medicine and Biology (OJEMB). Please notice that this is the template for TECHNOLOGY LETTERS. These are short manuscripts with primary focus on the development of new devices, methods, and technologies. The material must be organized in a main manuscript body and a section entitled Supplementary Materials. The main manuscript body should be limited to 1,000 words and the material should be organized in the following sections (in the order shown here): Introduction, Materials and Methods, Results, Discussion, Conclusions. Authors can include in the main manuscript body up to 3 display items (i.e. figures and tables). In addition, the main manuscript body should include an abstract and up to 20 references. The references are not considered in the 1,000-word count. The abstract should be organized in subsections as shown here (i.e., Goal, Methods, Results, Conclusions). An impact statement should be included. The impact statement should be a short paragraph of no more than 30 words. Additional material should be included in the Supplementary Materials section of the manuscript. The Supplementary Materials section should either be organized in sections as per the main manuscript body (using up to 4,000 words) or it should be used to provide readers with a set of up to 10 additional display items (i.e. figures and tables). \textit{Methods}: Use this document as a template if you are using Latex. Otherwise, use this document as an instruction set. \textit{Results}: Paper titles should be written in uppercase and lowercase letters as shown above. Do not cite references in the abstract. Do not delete the blank line immediately above the abstract; it sets the footnote at the bottom of this column. The abstract should not exceed 150 words. \textit{Conclusions}: Preparing carefully your manuscript will lead to enhanced readability.
\end{abstract}

@@

%TC:ignore
\begin{minipage}[t]{0.9\columnwidth}
  \textbf{\textit{Impact Statement}\textemdash{}Characterization of tinnitus can inform tinnitus treatment by allowing for individualized sound therapies, leading to better outcomes for patients suffering from the cognitive and psychological effects of tinnitus.}\\
\\
\end{minipage}
%TC:endignore

\section{Introduction}
% INTRO PARAGRAPH 1/3
\IEEEPARstart{T}{INNITUS}\textemdash{}the perception of sound
in the absence of any corresponding external stimulus\textemdash{}affects up to 50 million people in the U.S. \cite{bhattPrevalenceSeverityExposures2016,nationalinstituteondeafnessandothercommunicationdisordersQuickStatisticsHearing},
a third of whom experience functional cognitive impairment and substantial reduction in quality of life
\cite{ahmedImpactTinnitusPerception2017,nondahlImpactTinnitusQuality2007,tegg-quinnImpactTinnitusCognition2016}.
Clinical guidelines for tinnitus management frequently involve targeted exposure to tinnitus-like sounds, either as a form of habituation therapy (termed \textit{sound therapy}), or using Cognitive Behavioral Therapy \cite{tunkelClinicalPracticeGuideline2014}.
Critically, sound therapy is shown to improve with knowledge of the patient's internal tinnitus experience \cite{davisNeuromonicsTinnitusTreatment2007,landgrebeMethodologicalAspectsClinical2012,nickelOutcomeResearchMusic2005,okamotoListeningTailormadeNotched2010,schaetteAcousticStimulationTreatments2010,steinInhibitioninducedPlasticityTinnitus2015,tassCounteractingTinnitusAcoustic2012,wangStateArtSound2020}.
However, existing strategies for characterizing tinnitus sounds, such as Pitch Matching (PM) methods, are best suited for patients whose tinnitus resembles pure tones (\eg ringing), which may represent only $20-50\%$ of patients \cite{turnerjrAuditoryDysfunctionTinnitus1990,henryAuditoryTestResult2009,ukaegbeTinnitusItsEffect2017}.
There is a pressing need for more robust characterization methods \cite{henryTinnitusEpidemiologicPerspective2020,henryMeasurementTinnitus2016,norenaPsychoacousticCharacterizationTinnitus2002,robertsResidualInhibitionFunctions2006}.

% INTRO PARAGRAPH 2/3
Tinnitus sounds are presumed to be complex and heterogeneous
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,henryTinnitusEpidemiologicPerspective2020,meikleTinnitusArchiveArchive2004},
although few specifics have been firmly established.
Therefore, we base our approach on \textit{reverse correlation} (RC),
an established behavioral method
\cite{ahumadaStimulusFeaturesSignal1971,gosselinSuperstitiousPerceptionsReveal2003,brimijoinInternalRepresentationVowel2013}
for probing internal representations of visual and auditory percepts, unconstrained by prior knowledge.
RC asks subjects to render subjective judgments over random stimuli, 
from which unbiased estimates of the percepts can be produced.
RC is closely related to Wiener theory \cite{wienerNonlinearProblemsRandom1966},
which has inspired ``white-noise'' approaches to system characterization in physiology
\cite{ringachReverseCorrelationNeurophysiology2004,marmarelisWhiteNoiseMethodSystem1978} and engineering \cite{ljungMeasureLackFit1978}.

% INTRO PARAGRAPH 3/3
Here, we validate RC as a method for characterizing all component
frequencies of non-tonal tinnitus sounds
by estimating the full psychoacoustic tinnitus spectrum (PTS).
To that end, we asked normal-hearing participants to complete an augmented RC experiment, comparing random stimuli to a target sound.
The estimated PTS were subsequently validated against the targets.
Our results demonstrate, for the first time, that tinnitus-like sounds with complex spectra can be accurately estimated using RC.

@@

Stimuli were composed of $b=8$ Mel-spaced frequency bins,
which divide the frequency spectrum between $[100,~13,000]$
Hz into continuous segments of equal amplitude (\ie ``rectangular'' bins).
Reconstruction detail increases with $b$, but $b=8$ provides a good approximation to the chosen target sounds (see Fig. \ref{fig:spectra}).

For each stimulus, $[2,7]$ bins were randomly ``filled'' with power $0$ dB. ``Unfilled'' bins were assigned a power of $-100$ dB. 
Assigning random phase, the inverse Fourier transform of the spectrum yields a $500$ ms stimulus waveform.\footnote{Software for the experiments and analysis was written in MATLAB and is freely available at \protect\url{https://github.com/alec-hoyland/tinnitus-reconstruction/}}

@@

Two spectrally complex and complimentary target sounds (``buzzing'' and ``roaring'') were downloaded from the American Tinnitus Association \cite{MeasuringTinnitusAmerican2022} and truncated to $500$ ms in duration.

@@

10 normal-hearing subjects listened to the same target sound followed by a random stimulus for each trial (Fig. \ref{fig:experimentdiagram}).
The task is designed to mimic how a tinnitus patient might compare random stimuli to their tinnitus.

Subjects were instructed to respond ``yes'' to any stimuli that sounded similar to the target, and ``no'' otherwise.
Subjects completed $200$ trials for both target sounds within $10-20$ minutes.

@@

A subject performing $n$ RC trials with $b$ frequency bins produces a stimulus matrix $\Psi \in \mathbb{R}^{b \times n}$
and a response vector $y \in \{1,-1\}^n$,
where $1$ corresponds to a ``yes'' response and $-1$ to a ``no.'' RC classically assumes the subject response model:

\begin{equation}
  y = sign(\Psi^\mathrm{T} x), 
\end{equation}

where $x \in \mathbb{R}^\mathrm{b}$ is the subject's internal representation of interest (\eg their tinnitus percept).
This model can be inverted, yielding the following formula:

\begin{equation}
  \hat{x} = \frac{1}{n} \Psi y
  \label{eq:linreg}
\end{equation}

This formula is a restricted form of the Normal equation, under
the assumption that the stimulus dimensions are uncorrelated \cite{gosselinSuperstitiousPerceptionsReveal2003}.
Intuitively, $\hat{x}$ is the sum of all stimuli eliciting a ``yes'' response, after subtracting all stimuli eliciting a ``no.''

@@

The experimental paradigm allows for direct validation of the reconstructions $\hat{x}_{buzzing}$ and $\hat{x}_{roaring}$.
We represent the spectra of the target sounds as vectors $s_{buzzing} \in \mathbb{R}^\mathrm{b}$ and $s_{roaring} \in \mathbb{R}^\mathrm{b}$ using the same frequency bins as in stimulus generation with power equal to the mean power of frequencies within that bin.
Pearson's $r$ between $s_{buzzing}$ and $s_{roaring}$ and their corresponding reconstructions represents reconstruction accuracy.
One-sample t-tests were performed on the mean Fisher-transformed Pearson's $r$ values across subjects to assess significant differences from zero.

@@

To establish bounds on human performance,
additional experiments were run with two simulated subjects who give either \textit{ideal} or \textit{random} responses.
Each experiment ran for $n=200$ trials and was repeated $1000$ times.

@@

% NB NOTE: What is s??
for $i \in 1, ..., n$, where $Q(x, y)$ is the quantile function for $x \in [0, 1]$ of the similarity calculation $\Psi s$, and
$\Psi_i$ is the $i^\mathrm{th}$ column of $\Psi$.
The ideal subject has precise knowledge of every stimulus and responds mirroring Eq. \ref{eq:linreg}.

The \textit{random} subject responds with uniformly random probability.

@@

Figure \ref{fig:reconstructions} shows the distribution of Pearson's $r$ for human, ideal, and random subject responses.
Human accuracies were generally higher than random,
with some approaching ideal.
Mean random accuracy was $0$ in all conditions,
while mean human accuracy was significantly different from $0$ in all conditions: buzzing: $t(9) = 5.766$, $p < 0.001$; roaring: $t(9) = 5.76$, $p < 0.001$; combined: $t(19) = 7.542$, $p < 0.001$.
Greater variability in buzzing results may stem from the target spectrum containing significant high-frequency content, increasing the difficulty for subjects with minor high-frequency hearing loss.
Figure \ref{fig:spectra} plots the most accurate human reconstructions over the target sound spectra.

@@

Reconstruction accuracies observed here are well below the simulated ideal,
which may be attributed to noisy responses universally observed in applications of RC.
This may be mitigated by optimizing the experimental protocol, stimulus generation, and reconstruction method.
For example, recent approaches to improving RC reconstruction methods can boost efficiency, noise robustness and overall accuracy \cite{roopCompressiveSensingApproach2021,comptonStimulusWhiteningImproves2022a}.

Nevertheless, our results show that RC can accurately reconstruct the frequency spectrum of complex, non-tonal tinnitus-like sounds.
This indicates that RC may also be useful for developing a clinical assay to characterize a wider variety of tinnitus percepts than currently possible.
Additionally, subjects completed the required number of trials within 20 minutes,
indicating that this procedure could be conducted in a single clinical visit.
Future work will focus on validating this approach in patients suffering from tinnitus.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
% TC:ignore
\begin{abstract}
  \textit{Goal:} This study validates an approach to characterizing the sounds experienced by tinnitus patients via reverse correlation, with potential for characterizing a wider range of sounds than currently possible.
  \textit{Methods:} Ten normal-hearing subjects assessed the subjective similarity of random auditory stimuli and target tinnitus-like sounds (“buzzing” and “roaring”). Reconstructions of the targets were obtained by regressing subject responses on the stimuli, and were compared for accuracy to the frequency spectra of the targets using Pearson's $r$.
  \textit{Results:} Reconstruction accuracy was significantly higher than chance across subjects: buzzing ($M=0.53$, $SD=0.27$): $t(9)=5.766$, $p<0.001$; roaring ($M=0.57$, $SD=0.30$): $t(9)=5.76$, $p<0.001$.
  \textit{Conclusion:} Reverse correlation can accurately reconstruct non-tonal tinnitus-like sounds in normal-hearing subjects,
  indicating its potential for characterizing the sounds experienced by patients with non-tonal tinnitus.
% The purpose of this document is to illustrate how one should prepare manuscripts for submission to the IEEE Open Journal of Engineering in Medicine and Biology (OJEMB). Please notice that this is the template for TECHNOLOGY LETTERS. These are short manuscripts with primary focus on the development of new devices, methods, and technologies. The material must be organized in a main manuscript body and a section entitled Supplementary Materials. The main manuscript body should be limited to 1,000 words and the material should be organized in the following sections (in the order shown here): Introduction, Materials and Methods, Results, Discussion, Conclusions. Authors can include in the main manuscript body up to 3 display items (i.e. figures and tables). In addition, the main manuscript body should include an abstract and up to 20 references. The references are not considered in the 1,000-word count. The abstract should be organized in subsections as shown here (i.e., Goal, Methods, Results, Conclusions). An impact statement should be included. The impact statement should be a short paragraph of no more than 30 words. Additional material should be included in the Supplementary Materials section of the manuscript. The Supplementary Materials section should either be organized in sections as per the main manuscript body (using up to 4,000 words) or it should be used to provide readers with a set of up to 10 additional display items (i.e. figures and tables). \textit{Methods}: Use this document as a template if you are using Latex. Otherwise, use this document as an instruction set. \textit{Results}: Paper titles should be written in uppercase and lowercase letters as shown above. Do not cite references in the abstract. Do not delete the blank line immediately above the abstract; it sets the footnote at the bottom of this column. The abstract should not exceed 150 words. \textit{Conclusions}: Preparing carefully your manuscript will lead to enhanced readability.
\end{abstract}

@@

%TC:ignore
\begin{minipage}[t]{0.9\columnwidth}
  \textbf{\textit{Impact Statement}\textemdash{}Characterization of tinnitus sounds can inform treatment by facilitating individualized sound therapies, leading to better outcomes for patients suffering from the cognitive and psychological effects of tinnitus.}\\
\\
\end{minipage}
%TC:endignore

\section{Introduction}
% INTRO PARAGRAPH 1/3
\IEEEPARstart{T}{INNITUS}\textemdash{}the perception of sound
in the absence of any corresponding external stimulus\textemdash{}affects up to 50 million people in the U.S. \cite{bhattPrevalenceSeverityExposures2016,nationalinstituteondeafnessandothercommunicationdisordersQuickStatisticsHearing},
a third of whom experience functional cognitive impairment and diminished quality of life
\cite{ahmedImpactTinnitusPerception2017,nondahlImpactTinnitusQuality2007,tegg-quinnImpactTinnitusCognition2016}.
Clinical guidelines for tinnitus management involve targeted exposure to tinnitus-like sounds, for habituation therapy (termed \textit{sound therapy}), or as part of Cognitive Behavioral Therapy \cite{tunkelClinicalPracticeGuideline2014}.
Critically, sound therapy outcomes improve when closely informed by the patient's internal tinnitus experience \cite{davisNeuromonicsTinnitusTreatment2007,landgrebeMethodologicalAspectsClinical2012,nickelOutcomeResearchMusic2005,okamotoListeningTailormadeNotched2010,schaetteAcousticStimulationTreatments2010,steinInhibitioninducedPlasticityTinnitus2015,tassCounteractingTinnitusAcoustic2012,wangStateArtSound2020}.
However, existing strategies for characterizing tinnitus sounds, such as Pitch Matching (PM), are best suited for patients whose tinnitus resembles pure tones (\eg ringing), which may represent only $20-50\%$ of patients \cite{turnerjrAuditoryDysfunctionTinnitus1990,henryAuditoryTestResult2009,ukaegbeTinnitusItsEffect2017}.
There is a need for methods to characterize nontonal (\eg buzzing, roaring) tinnitus \cite{henryTinnitusEpidemiologicPerspective2020,henryMeasurementTinnitus2016,norenaPsychoacousticCharacterizationTinnitus2002,robertsResidualInhibitionFunctions2006}.

% INTRO PARAGRAPH 2/3
Nontonal tinnitus sounds are presumed to be complex and heterogeneous
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,henryTinnitusEpidemiologicPerspective2020,meikleTinnitusArchiveArchive2004},
although few characteristics have been firmly established.
Therefore, we base our approach on \textit{reverse correlation} (RC),
an established behavioral method
\cite{ahumadaStimulusFeaturesSignal1971,gosselinSuperstitiousPerceptionsReveal2003,brimijoinInternalRepresentationVowel2013}
for estimating internal perceptual representations that is unconstrained by prior knowledge about the representations themselves.
RC asks subjects to render subjective judgments over random stimuli, being closely related to Wiener theory \cite{wienerNonlinearProblemsRandom1966},
which has inspired ``white-noise'' approaches to system characterization in physiology
\cite{ringachReverseCorrelationNeurophysiology2004,marmarelisWhiteNoiseMethodSystem1978} and engineering \cite{ljungMeasureLackFit1978}.

% INTRO PARAGRAPH 3/3
Here, we validate RC as a method for characterizing a more complete psychoacoustic tinnitus spectrum (PTS) for tinnitus sounds.
To that end, normal-hearing participants completed an augmented RC experiment, comparing random stimuli to a target tinnitus-like sound.
The estimated PTS was subsequently validated against the targets.
Our results demonstrate, for the first time, that tinnitus-like sounds with complex spectra can be accurately estimated using RC.

@@

Stimuli were composed of $b=8$ Mel-spaced frequency bins,
which divide the frequency spectrum between $[100,~13,000]$
Hz into contiguous segments of equal amplitude (\ie ``rectangular'' bins).
Reconstruction detail increases with $b$, but $b=8$ provides a good approximation to the chosen target sounds (see Fig. \ref{fig:spectra}).

For each stimulus, $[2,7]$ bins were randomly ``filled'' with power $0$ dB. ``Unfilled'' bins were assigned $-100$ dB. All frequencies were assigned random phase.
Inverse Fourier transform of the constructed spectrum yields a $500$ ms stimulus waveform.\footnote{Software for the experiments and analysis was written in MATLAB and is freely available at \protect\url{https://github.com/alec-hoyland/tinnitus-reconstruction/}}

@@

Two spectrally complex and complimentary target sounds (``buzzing'' and ``roaring'') were downloaded from the American Tinnitus Association \cite{MeasuringTinnitusAmerican2022} and truncated to $500$ ms in duration (Fig. \ref{fig:spectra}).

@@

Ten ($n=10$) normal-hearing subjects listened to $A-X$ trials containing a target sound ($A$) followed by a stimulus ($X$). $X$ was randomly generated for each trial, while $A$ remained the same within a block of $100$ trials (Fig. \ref{fig:experimentdiagram}). Subjects completed two (2) blocks per target sound. Subjects were told that some stimuli had $A$ embedded in them, and were instructed to respond ``yes'' to such stimuli, otherwise ``no''. Subjects listened over earphones at a self-determined comfortable level. Procedures were approved by the UMass IRB.

@@

A subject performing $p$ RC trials with $b$ frequency bins produces a stimulus matrix $\Psi \in \mathbb{R}^{p \times b}$
and a response vector $y \in \{1,-1\}^n$,
where $1$ corresponds to a ``yes'' response and $-1$ to a ``no.'' RC classically assumes the subject response model:

\begin{equation}
  y = sign(\Psi^\mathrm{T} x),
  \label{eq:subjresp} 
\end{equation}

where $x \in \mathbb{R}^\mathrm{b}$ is the subject's internal representation of interest (\eg of their tinnitus).
Inverting this model yields:

\begin{equation}
  \hat{x} = \frac{1}{n} \Psi y
  \label{eq:linreg}
\end{equation}
which is a restricted form of the Normal equation under
the assumption that the stimulus dimensions are uncorrelated \cite{gosselinSuperstitiousPerceptionsReveal2003}.

@@

The experimental paradigm allows for direct validation of the reconstructions $\hat{x}_{buzzing}$ and $\hat{x}_{roaring}$.
We represent the spectra of the target sounds as vectors $s_{buzzing} \in \mathbb{R}^\mathrm{b}$ and $s_{roaring} \in \mathbb{R}^\mathrm{b}$ using the same frequency bins as the stimulus with power equal to the mean power at frequencies within that bin.
Pearson's $r$ between $s_{buzzing}$ and $s_{roaring}$ and their corresponding reconstructions quantifies reconstruction accuracy.
One-sample t-tests were performed on the mean Fisher-transformed Pearson's $r$ values across subjects to assess significant differences from zero.

@@

To establish bounds on human performance,
additional experiments were run with two simulated subjects who give either \textit{ideal} or \textit{random} responses.
Each experiment ran for $p=200$ trials and was repeated $1000$ times.

@@

% NB NOTE: What is s?? ACL: it is the target sound vector
for $i \in 1, ..., p$, where $Q(x, y)$ is the quantile function for $x \in [0, 1]$ of the similarity calculation $\Psi s$, and
$\Psi_i$ is the $i^\mathrm{th}$ column of $\Psi$.
Thus, the ideal subject has precise knowledge of every stimulus and responds according to Eq. \ref{eq:subjresp}. The \textit{random} subject responds $y_i \in \{1,-1\}^n$ with uniform random probability, thus ignoring the stimulus entirely.

@@

Figure \ref{fig:reconstructions} shows the distribution of Pearson's $r$ for human, ideal, and random subject responses.
Human accuracies were generally higher than random,
with some approaching \textit{ideal}. 
Mean human accuracy was significantly different from $0$ (i.e., mean \textit{random} accuracy) in all conditions: buzzing: $t(9) = 5.766$, $p < 0.001$; roaring: $t(9) = 5.76$, $p < 0.001$; combined: $t(19) = 7.542$, $p < 0.001$.
Figure \ref{fig:spectra} plots the most accurate human reconstructions over the target sound spectra.

@@

Our results show that RC can accurately reconstruct the PTS of complex, non-tonal, tinnitus-like sounds, capturing the most salient features of the target.
Reconstruction accuracies observed here are below the simulated ideal, which may be attributed to noisy responses universally observed in applications of RC, and which may be mitigated by further optimizing the experimental protocol, stimulus generation, and reconstruction method.
For example, recent approaches to improving RC reconstruction methods can boost efficiency, noise robustness and overall accuracy \cite{roopCompressiveSensingApproach2021,comptonStimulusWhiteningImproves2022a}.
Subjects completed the required number of trials within 10 minutes,
indicating that this procedure could be conducted in a single clinical visit.
RC may therefore be useful as the basis for a clinical assay to characterize a wider variety of tinnitus percepts than currently possible.
Future work will focus on validating this approach in patients suffering from tinnitus.
